---
title: "sok_1005_arbeidskrav4"
format: html
editor: visual
---

## Task 1

```{r}
#Clear list
rm(list = ls())
```

```{r}
#Loading packages
library(jsonlite)
library(tidyverse)
library(scales)
library(ggrepel)
library(ggplot2)
```

```{r}
#Json-link
df_json <- fromJSON("https://static01.nyt.com/newsgraphics/2021/12/20/us-coronavirus-deaths-2021/ff0adde21623e111d8ce103fedecf7ffc7906264/scatter.json")

#Adding statecodes
df_json <- df_json %>%
  mutate(state_name = state.abb[match(name, state.name)])

#Fixing Na
df_json[9, 7] = "D.C"
```

```{r}
#Plotting figure
figure_json <- ggplot(df_json, aes(x = fully_vaccinated_pct_of_pop*100, 
                    y = deaths_per_100k,
                    label = state_name)) +
  theme_bw() +
  geom_point(size = 1) +
  geom_text_repel(size = 3) +
  labs(x = "Percentage of fully vaccinated people",
       y = "",
       title = "Covid-19 deaths since universal adult vaccine eligibility compared with
       vaccination rates",
       subtitle = "20 avg. monthly deaths per 100,000",
       caption = "Source: New York Times Database") +
  scale_x_continuous(labels = percent_format(scale = 1),
                     breaks = seq(45,80,5),
                     limits = c(45,80),
                     expand = c(0, 0)) +
  scale_y_continuous(breaks = seq(0,20,5),
                     limits = c(0,20),
                     expand = c(0,0)) +
  annotate("text",
           x=56,y=18,
           label = "Lower vaccination rate,
           higher death rate",
           size = 3) +
    geom_segment(aes(x=52.5, y=18, xend=51,yend=19),
               arrow = arrow(length = unit(.2, 'cm')), lwd = .5) +
  
    annotate("text",
           x=75,y=10,
           label = "Higher vaccination rate,
           lower death rate",
           size = 3) +
  geom_segment(aes(x=77, y=8.8, xend=79,yend=7.8),
               arrow = arrow(length = unit(.2, 'cm')), lwd = .5) +
  
  theme(panel.border = element_rect(linetype = 3),
        panel.grid = element_line(linetype = "dotted"))
        
figure_json
```

## Task 2

```{r}
lm(deaths_per_100k ~ fully_vaccinated_pct_of_pop, data = df_json)
```

```{r}
figure_lm <- figure_json + geom_smooth(method = "lm", se = FALSE)

figure_lm
```

The linear model shows that there is a clear trend between fully vaccinated people and deaths per 100k, the higher vaccination rate, the lower death rate. Furthermore, the coefficients also tells us the relationship between these two variables. It is important to take into consideration that I have multiplied the percentage of fully vaccinated people. So the coefficient needs to be divided by 100. Which means that per percent of fully vaccinated people there will be approximately 0,37 less deaths per 100,000. Or per 10,000,000 there will be 37 less deaths.
